Placing the spotted T Tauri star LkCa 4 on an HR diagram

### 2016-08-15 document/ms.tex
Add lots of text for the discussion, related to issue #36

\subsection{Longitudinal / latitudinal spot distributions and phase}

\subsection{Addressing limitations of assumptions in the methodology}

\emph{Number of photospheric components-}  Our full spectrum fitting method assumed \emph{only} two temperature components present in the photosphere.  In reality, starspots are probably described by a range of effective temperatures like umbra, penumbra, and plages on the sun's surface.  A model possessing three unique temperatures and two unique fill factors might better approximate the minor constituents of the emergent spectrum than our two temperature model can.  Computational infeasibility limited our ability to rerun new multi-temperature models and compute model comparison diagnostics like AIC, BIC, or cross-validation \citep{ivezic14}.  Such investigations could be considered in future works, but we suspect systematic errors in the models would ultimately limit the interpretability of models with more than two temperature components.  High-precision polychromatic planetary transit photometry could offer an avenue to measure the morphologies and temperatures of starspots for the subset of transiting exoplanet host stars.
% The fact that we get consistent estimates for Teff tells us something.
% The fits are pretty good, also tells us something.
% Coarsely sampled models + spectral emulator acts as a smoother in Teff, as does the GP global covariance kernel.
%(cite Charlotte Norris CS19 poster?)

\emph{Suitability of pre-computed \PHOENIX\ model spectra, model ranges, $\A \neq 0$, and magnetic fields-} We have assumed that the \PHOENIX\ synthetic spectra are a good representation of our observed IGRINS spectrum of \name, a young star likely to possess a large magnetic field.  Cool sunspots are associated with heightened magnetic fields on the sun.  Magnetic fields will have two effects on the emergent spectrum- line broadening attributable to the Zeeman effect, and pressure broadening attributable to seeing deeper into deeper layers in the photosphere compared to the surrounding hotter photosphere.  The former effect will differentially impact lines with high Lande-g factors (cite XX, fix the g formatting).  The latter effect could masquerade as an apparent shift in surface gravity, as seen in mixture models of spotted stars with empirical templates composed of dwarfs and giants (cite XX Saar and collaborators).  Our tacit assumption has been that these effects are 1) relatively small compared to the gross appearance and disappearance of spectral lines at disparate temperatures and 2) there exist many orders and lines that are not affected at all.  We see tentative evidence for physics outside of the pre-computed model grids in the poor fits to the observed spectra in $K-$band.  Strong metal lines demonstrate large departures in their best fit $\Z$ due to Zeeman broadening.  No pre-computed model grids including a range of surface magnetic fields are available at the wavelength range and spectral sampling needed to approach these questions within this spectral inference framework.
% Note that these two effects would differently affect the spectra

%Our Gaussian process inference framework yields some resilience to the former problem.

\subsection{Umbra versus penumbra}
Sunspots consist of cool penumbra enveloping much cooler and smaller umbra with temperature differences of X and Y respectively \todo{cite}.  We must be seeing the umbra of the star.

\subsection{Interpretation of the large filling factor of cool starspots}

\subsection{Is \name an extreme source?}

\subsection{Starspots as a source of RV Jitter}
Starspots cause significant jitter in radial velocity surveys for young planets \citep[e.g.][]{donati14, robertson14}.  \todo{Potentially expand on this topic, but not critical.}

This observation has motivated precision radial velocity (RV) planet searches in the infrared, where the modulation of RV signals is lessened \citep{prato08}.

@@

Table \ref{tbl_order_results} lists all the IGRINS spectral order results.
\comm{It's not clear if we need this table.  If we keep it, we should update it with a reliability index or note about which orders were used in deriving the final values.  A better table might be a list of new spectral features attributable to the spot.}
\input{tables/tbl_order_results.tex}

>>>>

\subsection{Addressing limitations of assumptions in the spectral inference methodology}

The spectral inference method assumed only two temperature components present in the photosphere.  In reality, starspots are probably described by a range of effective temperatures like umbra, penumbra, and plages on the sun's surface.  A model possessing three unique temperatures and two unique fill factors might better approximate the minor constituents of the emergent spectrum than our two temperature model can, but would always fit the data better.  The fit-quality of the existing two-temperature fits seems fairly strong through much of $H-$band and many $K-$band orders, so the contribution of minor photospheric constituents must be relatively low.  Interestingly, $K-$band has several spectral orders that converge towards hybrid temperatures in the vicinity of $\sim3500$ K.  Most of these orders were flagged as erroneous due to strong metal line spectral outliers or uncorrected calibration artifacts.  But some orders just showed mediocre fits where no permutation of hot and cool two-temperature models could describe all of the variance in the spectrum.  These $K-$band orders could be examined for a penumbra-like temperature component.  We suspect systematic errors in the models would ultimately limit the interpretability of models with more than two temperature components.  High-precision polychromatic planetary transit photometry could offer an avenue to measure the morphologies and temperatures of starspots for the subset of transiting exoplanet host stars.

\emph{Suitability of pre-computed \PHOENIX\ model spectra, model ranges, $\A \neq 0$, and magnetic fields-} We have assumed that the \PHOENIX\ synthetic spectra are a good representation of the observed IGRINS spectrum of \name, a young star likely to possess a large magnetic field.  Cool sunspots are associated with heightened magnetic fields on the sun.  Magnetic fields will have two effects on the emergent spectrum- line broadening attributable to the Zeeman effect, and pressure broadening attributable to seeing deeper into deeper layers in the photosphere compared to the surrounding hotter photosphere.  The former effect will differentially impact lines with high magnetic sensitivity \citep{2013AJ....146...51D}.  The latter effect could masquerade as an apparent shift in surface gravity, as seen in mixture models of spotted stars with empirical templates composed of dwarfs and giants \citep{oneal96}.  Our tacit assumptions have been that these effects are 1) relatively small compared to the gross appearance and disappearance of spectral lines at disparate temperatures, 2) there exist many orders and lines that are not affected at all, and 3) the Gaussian process noise model allows some resilience against unknown residual spectrum correlations.  We see tentative evidence for physics outside of the pre-computed model grids in the poor fits to the observed spectra in $K-$band.  Strong metal lines demonstrate large departures in their best fit $\Z$ due to Zeeman broadening.  No pre-computed model grids including a range of surface magnetic fields are available at the wavelength range and spectral sampling needed to approach these questions within this spectral inference framework.

\subsection{Interpretation of the large filling factor of cool starspots}
We have shown that LkCa 4 exposes between 67\% and 83\% cool photosphere towards our viewing direction.  This exceptionally large fraction of surface area covered by spots defies the conventional wisdom that starspots should not cover more than 50\% of a stellar surface, since the minority constituent of temperature is usually called the spot.  The evidence for this derived fill factor has been shown in the SED, high-resolution spectral inference, direct spectral detection, and TiO variability modelling.  Perhaps the strongest evidence for the large fill factor of cool photosphere is the large amplitude of $V-$band modulation, which cannot be explained with small spots, even if an effective temperature of absolute zero is assigned to the cool regions.

One conceivable interpretation of the 67\% to 83\% coverage fraction of cool photosphere is a stellar axis inclined towards our observing direction, with a large circum-polar spot covering much of the observable hemisphere.  The ambient hot photosphere would then be close to the equator, which could additionally be limb-darkened due to project effects.  Local fingers of cool spots would reached down to near the equator, entering into and out-of the field of view and giving rise to the large amplitude photometric variation seen on LkCa 4.  \todo{Is this picture consistent with Donati et al. 2014?}.

\subsection{Is \name an extreme source?}
LkCa 4 is an extreme source by several metrics.  Mainly, its large amplitude of cyclical optical variability exceeds all other counterparts from long-term monitoring of WTTSs \citep{grankin08}.  LkCa 4 has about 40 times the photometric amplitude of the typical young star towards Ophiuchus or Upper Sco (Gully-Santiago et al. in prep.).

\subsection{Evidence for intra-spectrum RV jitter}
Starspots cause significant jitter in radial velocity surveys for young planets \citep[e.g.][]{donati14, robertson14}.  The two-temperature model assumed that the hot and cool components shared the same $\vsini$ and $v_z$.  This assumption is accurate for either homogeneously distributed spots, azimuthally symmetric bands of spots, or polar spots.  However, LkCa 4 has clear optical line-profile variations \citep{nguyen12, donati14}.  Interestingly, we should be able to detect anticorrelation between lines attributable to hot photosphere, and lines attributable to cool photosphere, since there is a zero-sum competition for solid-angle on the recessional and advancing stellar limbs.  We found tentative, albeit weak, evidence for such anticorrelation in $v_z$ across spectral orders with dominated by cool photosphere and those dominated by hot photosphere.  For example, the cool-photosphere dominstated spectral orders $o = 100, 102, and 104$ show the largest median $v_z\sim14.5-17.0$ km/s, while the other orders that possess a mix of primarily hot- and some cool- photosphere features demonstrate $v_z\sim9-13$ km/s.

Radial velocity (RV) planet searches in the infrared--where the modulation of RV signals is lessened \citep{prato08}-- could potentially reduce RV jitter by characterizing and constraining the effect of direct emission from starspots.

\subsection{Comparison with APOGEE/IN-SYNC}
The INfrared Spectra of Young Nebulous Clusters (IN-SYNC) project \citep{cottaar14} employs APOGEE, which has a $15100-17000 \;\AA$ bandwidth.  The long-wavelength cutoff prevents APOGEE from detecting the most conspicuous cool star emission we detected in IGRINS spectral orders $o = 100, 101, 104$ at $17075-17990 \;\AA$.  These spectral orders can be thought of as a ``sweet spot'' where the cool star spectrum is putting out ample flux, but still far-enough away from the $K-$band where problems like magnetic field-sensitive lines, and enhanced telluric absorption complicate the spectra.  Additionally, veiling from mid-IR excess in classical T-Tauri stars will be lessened in $H-band$.

\subsection{Addressing binarity}
In principle, unresolved eclipsing binaries could mimic some of the data shown here.  However, unresolved such a system would have had to evade detection in high-resolution imaging \citep{kraus11,daemgen15}...

@@

Table \ref{tbl_order_results} lists all the IGRINS spectral order results.
\comm{Update this with ``used/unused'' label and most recent values.}
\input{tables/tbl_order_results.tex}
